Manuscript

### 2025-03-27 Manuscript_MD_ScoreReLiability.qmd
minor fixed text arrange Results

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. For attempts
to estimate a phenomenon’s replicability or to predict whether a future
replication will replicate successfully, heterogeneity is a crucial
parameter that needs to be assessed. In the meta-analytic context,
heterogeneity describes the variation in effect sizes, free of sampling
error, if these are not constant across replications. Therefore, the
presence of heterogeneity implies that some replications of a single
phenomenon may be successful, while others are not. If heterogeneity
grows, meaning the phenomenon’s effect size varies more strongly for
unexplained reasons, the probability of observing an effect size around
zero or even in the negative space grows larger as well. If we know the
mean size and heterogeneity of a phenomenon’s effect size, we would
theoretically be able to establish an expected replication rate
(reference).

@@

Large-scale attempts of direct replications, using identical protocols,
such as the Many Labs studies or Registered Replication Reports
(references), for the first time allow researchers to estimate
heterogeneity undistorted by selection processes induced by publication
bias. In re-analyses of these studies, Olsson-Collentine et al. (2020)
identify a positive correlation between a phenomenon’s effect size and
its heterogeneity. Similarly, van Erp et al. (2017) or Stanley et al.
(2018) estimate strong degrees of heterogeneity across conceptual
replications in Psychology. In a separate re-analysis of large scale
direct replications, Renkewitz et al. (2025) identify substantial
heterogeneity in almost all projects where a non-zero effect could be
identified. This aligns with the correlation found by Olsson-Collentine
et al. (2020).

@@

We have collected a large number of openly available multi-site direct
replications on psychological phenomena (Fünderich et al. 2024). The
data-sets of about 50 phenomena , largely stemming from the efforts of
the Many Labs studies, Registered Replication Reports or the
Psychological Science Accelerator, are made openly available in a
standardised format at \[osf.io\], the DRIPHT repository.

@@

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects was made publicly available
when the DRIPHT repository was set up (references). The Registered
Replication Reports are similar collaborative efforts, with the sole
distinction that for each report a single phenomenon was replicated
several times. The Registered Reports 3-10 were added to the DRIPHT
repository, as they employed experimental designs with at least two
groups. Lastly, some projects from the Psychological Science Accelerator
were added to the DRIPHT repository. Different to Many Labs or
Registered Replication Projects, these collaborative efforts, while also
distributed across the globe, do not formally attempt to perform direct
replications, but focus on original research or conceptual replications.
However, as for each project, across all sites the same protocol is
used, and the subjects are distributed across different countries, a
data structure similar to that of a multi-site replication study
emerges. What makes all these collaborative efforts valuable for this
manuscript is the fact that all phenomena collected in the
DRIPHT-repository are making use of two-group designs and are therefore
easily assessed using standardised effect sizes.

@@

@tbl-meanES demonstrates that what we observed in Figure 1 holds across
all 21 phenomena. All meta-analytic effects are larger in their absolute
size after applying an attenuation correction procedure. Similarly, the
uncertainty quantified in the estimates’ standard error is larger.
Additionally, @tbl-meanES highlights which phenomena pass criterion (3)
– the effect size must be statistically significantly distinguishable
from zero. 12 phenomena pass this criterion, highlighted in boldface. It
may be remarked that the attenuation correction procedure, in this data,
appears to have no influence on whether a phenomenon passes the
criterion. The differences in p-value are rather small and lead to no
difference in conclusion, regarding a significance level of .05.

@@

### Results

@@

tab_hetES <- read.csv(here("Tables/Heterogeneity_ES.csv"))
tab_hetrel <- read.csv(here("Tables/Results_RMA_alpha.csv"))
tab_muES <- read.csv(here("Tables/Mean_ES.csv"))

@@

knitr::kable(tab_both,
             col.names = c("MASC", 
                           "$\\hat{\\mu}_{\\alpha}$ (95%-CI)", 
                           "$\\hat{\\tau}_{\\alpha}$", 
                           "QE (df)", "p", 
                           "$\\hat{\\tau}_{\\delta_0}$", 
                           "QE (df)", "p",
                           "$\\hat{\\tau}_{\\delta}$", 
                           "QE (df)", "p"),
             escape = FALSE)
# 
# tab_one <- data.frame(tab_muES,
#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
```

@@

## Alternative discussion of how score reliability affects ES heterogeneity

@@

### Misinterpreted equation in Hunter & Schmidt (XXXX)

@@

The metric $R_2$ on the other hand describes, in how far the attenuation
correction procedure has successfully reduced heterogeneity in score
variance $\tau^2_{\sigma^2_X}$. A value of 1 indicates that the
heterogeneity in true score variance $\tau^2_{\sigma^2_T}$, the score
variance after the attenuation correction, is essentially just as large
as the heterogeneity initially observed. Smaller values indicate how
much heterogeneity is "left", after applying an attenuation correction,
relative to the initial heterogeneity. This means that a value of .7
indicates that about 70% of heterogeneity in score variance remains,
even after correcting for differences in score reliability. As only 30%
of score variance heterogeneity could be removed, this would imply that
differences in error score variance $\sigma^2_E$ were responsible for
less than a third of the heterogeneity in score variances found. On the
contrary, more than two thirds of heterogeneity could be attributed to
actual differences in how the underlying true scores are distributed
across samples.

Concerning the re-arranged inequality from Appendix A, we identify one
crucial condition described in @eq-ineq_short.

@@

As long as $R_1^3$ is smaller or equal compared to $R_2$, @eq-ineq_init
is bound to hold true. This means that as long as the average score
reliability, to the power of 3, does not exceed the reduction in score
variance heterogeneity due to the attenuation correction, @eq-ineq_init
is not violated. For example, assuming the average score reliability is
about .8, as long as $R_2$ is at least .512 or larger, the inequality
holds true and ES heterogeneity grows larger as a consequence of the
attenuation correction procedure. In such cases, the claims made in W&D
and H&S could not hold.

$R_1$ essentially describes the reduction in mean score variance due to
the attenuation correction, and $R_2$ essentially described the
reduction in heterogeneity of score variance due to the attenuation
correction. Therefore, @eq-ineq_init states that for such a correction
procedure to truly reduce ES heterogeneity, the reduction in the
heterogeneity of score variance needs to be substantially larger than
the reduction in its mean values. The mean value reduction carries
stronger weight, as $R_1$ is taken to the power of 3 in @eq-ineq_init.

@@

@@

Estimation of $R_1$ and $R_2$ require estimates of meta-analytic mean
and heterogeneity of both the total score variance $\sigma^2_X$ and the
true score variance $\sigma^2_T$. Elsewhere, we proposed the method
boot-err (ref.) as an alternative method to test for and estimate the
differences in error score variance $\sigma^2_E$. In this text, we
deliver detailed instructions on how estimates for mean and
heterogeneity of total score variance, and the score variance components
can be derived. For the total score variance $\sigma^2_X$, all we need
to do is to appropriately transform the observed estimates using the
variance-stabilizing properties of the natural log. Subsequently, we use
bootstrapping to derive robust standard errors for these transformed
observations. Using a random-efffects meta-analysis, we derive estimates
of mean and heterogeneity, which need to be backtransformed to the
original scale. Appropriate equation to back-transform estimates can be
found in Appendix B.

@@

For one of the phenomena, however, we observe a large disparity compared
to the rest of the results. For the Shnabel_Willingness_Reconcile_RPP
phenomenon, $R_2$ is very low, below .25. This makes it even lower than
its $R_1^3 = .574$. This would imply that for this phenomenon, the
conditions for a reduction in ES heterogeneity would be met. However, in
@tbl-tauES, we find no such reduction. What becomes apparent, however,
is that for this phenomenon we are working with the low number of 9
administration sites. Estimation precision for heterogeneity in variance
components, even more so than for ES, largely depends on the sample
size. Beinhauer et al. (ref) come to the conclusion that at 10 items,
which the WIllingness to Reconcile scale consists of, a sample size
around 33 is required to achieve sufficient power for a statistical test
of score variance heterogeneity. Most likely, the low size led to
implausible estimates in both true and total score variance
heterogeneity and subsequently to an implausible estimate of $R_2$.
Meta-analytic estimates, especially those concerning score variance
components such as $R_2$, therefore should be interpreted with caution,
if the sample size was particularly low.

@@

Generally, @tbl-tauES implies that for almmost all measurements
concerning these phenomena, the heterogeneity in observed score
variances could not be substantially explained by differences in
measurement quality. Only for one phenomenon, where differences in score
variance could be found, $R_2$ was smaller than $R_1^3$. As discussed,
for this phenomenon, the number of administration sites was too low to
accurately estimate heterogeneity in score variance components. For all
remaining phenomena, $R_2$ was larger than $R_1$, implying that an
attenuation correction procedure could not reduce ES heterogeneity. The
results in @tbl-tauES agree with this conclusion, as no phenomena could
be identified where statistically significant ES heterogeneity was
accompanied by a reduction in differences due to the correction.

```{r, echo = FALSE}
#| label: tbl-R1R2
#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
knitr::kable(read.csv(here("Tables/Variances_analysis.csv")))
```

@@

In the examples we re-analysed throughout this text, as well as the
majority of research studies done in the behavioural sciences, no proper
sampling techniques are employed. Most samples where these instruments
are applied come to be through convenience sampling, where students
self-select into studies. Without a valid sampling scheme, we
essentially have have no basis to assume anything about the true score
variance. Most importantly, however, we have no reason to assume that
the true score variance could potentially be homogeneous or stable
across administration sites. Instead, it is very likely that the
convenience sampling procedures produce samples that vary extremely
strongly in how they are put together, leading to large degrees of true
score variance heterogeneity. At the same time, score reliability is a
low-hanging fruit for researchers to optimise most researchers are aware
and trained in assessing score reliability.

@@

If we, in the space of behavioural sciences, keep making use of
convenience samples simply because they are so convenient, we can not
expect anything to change. Most likely, ES heterogeneity will remain a
substantial factor hampering the interpretability of original and
replication studies. We could learn more about standardized effect sizes
and why our phenomena vary across settings, populations or continents by
not following the current status quo of convenience samples. Until this
status quo changes, however, we will keep wasting valuable resources by
attempts of discussing ES heterogeneity or assessing uninformative
differences in score reliability.

@@

The analytical arguments presented following Equation 17 rested on the
use of the delta-method, as Equation 17 contained parameters of the
score standard deviations’ distributions, while the $R_1$ and $R_2$
metrics were designed to discuss score variances. The delta-method rests
on a number of assumptions: (1) approximations derived from the
delta-method require for the transformed variable to follow a marginal
normal distribution (ref). Since the variable under discussion here is
the standard deviation, which is bounded to be larger than zero and has
a non-normal sampling distribution (ref), this assumption is most likely
violated; (2) the derivative of the transformation function needs to be
available. As the transformation function is essentially just the square
root, this assumption should hold up; (3) as the delta-method
approximates the parameters of the transformed variable using the
first-order Taylor-expansion, this implies that higher-order terms are
required to be negligible. In practice, this implies that the variance
of the untransformed variable needs to be small. However, as this entire
procedure discusses the influence of variance in standard deviations on
ES heterogeneity, introduced by differences in score reliability, this
assumption is probably violated as well. The violation of these
assumptions likely introduce a bias in estimates derived from these
approximations (ref). However, the inequality derived with use of the
delta-method is not used to actually derive any estimates, but to
discuss how an attenuation correction procedure can affect ES
heterogeneity. Therefore, we believe that the violation of these
assumptions is defensible in this case, as the analytical arguments
derived from the inequality should still hold up, nevertheless. \[add to
appendix?\] Additionally, we used the empirical examples available to
assess in how far estimates derived using the delta-method
approximations diverge from estimates derived without it. The
differences identified are rather miniscule at the third decimal,
reinforcing our belief that these assumption violations are defensible.

>>>>

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. For attempts
to estimate a phenomenon’s replicability or to predict whether a future
replication will replicate successfully, heterogeneity is a crucial
parameter that needs to be assessed. In the meta-analytic context,
heterogeneity describes the variation in effect sizes, free of sampling
error, if these are not constant across replications. Therefore, the
presence of heterogeneity implies that some replications of a single
phenomenon may be successful, while others are not. If heterogeneity
grows, meaning the phenomenon’s effect size varies more strongly for
unexplained reasons, the probability of observing an effect size around
zero or even in the negative space grows larger as well. If we know the
mean size and heterogeneity of a phenomenon’s effect, we would
theoretically be able to establish an expected replication rate
(reference).

@@

Large-scale attempts of direct replications, using identical protocols,
such as the Many Labs studies or Registered Replication Reports
(references), for the first time allow researchers to estimate
heterogeneity undistorted by selection processes induced by publication
bias. In re-analyses of these studies, Olsson-Collentine et al. (2020)
identify a positive correlation between a phenomenon’s effect size and
its heterogeneity. Similarly, van Erp et al. (2017) or Stanley et al.
(2018) estimate strong degrees of heterogeneity across conceptual
replications in Psychology. In a separate re-analysis of large scale
direct replications, Renkewitz et al. (in preparation) identify
substantial heterogeneity in almost all projects where a non-zero effect
could be identified. This aligns with the correlation found by
Olsson-Collentine et al. (2020).

@@

We have collected a large number of openly available multi-site direct
replications on psychological phenomena (Fünderich et al. 2024). The
data-sets of about 50 phenomena, largely stemming from the efforts of
the Many Labs studies, Registered Replication Reports or the
Psychological Science Accelerator, are made openly available in a
standardised format at \[osf.io\], the DRIPHT repository.

@@

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects was made publicly available
when the DRIPHT repository was set up (references). The Registered
Replication Reports are similar collaborative efforts, with the sole
distinction that for each report a single phenomenon was replicated
several times. The Registered Reports 3-10 were added to the DRIPHT
repository, as they employed experimental designs with at least two
groups. Lastly, some projects from the Psychological Science Accelerator
were added to the DRIPHT repository. Different to Many Labs or
Registered Replication Projects, these collaborative efforts, while also
distributed across the globe, do not formally attempt to perform direct
replications, but focus on original research or conceptual replications.
However, as for each project, across all sites the same protocol is used
and the subjects are distributed across different countries, a data
structure similar to that of a multi-site replication study emerges.
What makes all these collaborative efforts valuable for this manuscript
is the fact that all phenomena collected in the DRIPHT-repository are
making use of two-group designs and are therefore easily assessed using
standardised effect sizes.

@@

@@

### Results - Wich phenomena pass criterion 3)?

@tbl-meanES demonstrates that what we observed in Figure 1 holds across
all 21 phenomena. All meta-analytic effects are larger in their absolute
size after applying an attenuation correction procedure. Similarly, the
uncertainty quantified in the estimate's standard error is larger.
Additionally, @tbl-meanES highlights which phenomena pass criterion (3)
– the effect size must be statistically significantly distinguishable
from zero. 12 phenomena pass this criterion, highlighted in boldface. It
may be remarked that the attenuation correction procedure, in this data,
appears to have no influence on whether a phenomenon passes the
criterion. The differences in p-value are rather small and lead to no
difference in conclusion, regarding a significance level of .05.

### Results - Estimates of heterogeneity

@@

tab_hetES <- read.csv(here("Tables/Heterogeneity_ES.csv"))
tab_hetrel <- read.csv(here("Tables/Results_RMA_alpha.csv"))

@@

knitr::kable(tab_both,
             col.names = c("MASC", 
                           "$\\hat{\\mu}_{\\rho_{XX'}}$ (95%-CI)", 
                           "$\\hat{\\tau}_{\\rho_{XX'}}$", 
                           "QE (df)", "p", 
                           "$\\hat{\\tau}_{\\delta_0}$", 
                           "QE (df)", "p",
                           "$\\hat{\\tau}_{\\delta}$", 
                           "QE (df)", "p"),
             escape = FALSE)
# 
# tab_one <- data.frame(tab_muES,
#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
```

*Note:* $\hat{\mu}_{\rho_{XX'}}$ is the back-transformed meta-analytic
estimate of mean score reliability; $\tau{\mu}_{\rho_{XX'}}$ is the
back-transformed estimate of heterogeneity in score reliability; Note
that, since in a meta-analysis, we attempt to estimate mean and variance
of the parameter's true distribution, in this table
$\hat{\mu}_{\delta_0}$ represents the meta-analytic estimate of mean
uncorrected $d$ and $\hat{\tau}_{\delta_0}$ represents the estimate of
its heterogeneity; accordingly, $\hat{\mu}_{\delta}$ represents the
estimate of the meta-analytic mean of corrected $d_c$ and
$\hat{\tau}_{\delta}$ the estimate of its heterogeneity.

@@

## Relationship between score reliability and effect sizes in terms of heterogeneity

@@

### Hunter & Schmidt's misinterpreted equation

@@

The metric $R_2$ on the other hand describes in how far the attenuation
correction procedure has successfully reduced heterogeneity in score
variance $\tau^2_{\sigma^2_X}$. A value of 1 indicates that the
heterogeneity in true score variance $\sigma^2_T$ is essentially just as
large as the heterogeneity initially observed. Smaller values indicate
how much heterogeneity is "left", after applying an attenuation
correction, relative to the initial heterogeneity. This means that a
value of .7 indicates that about 70% of heterogeneity in score variance
remains, even after correcting for differences in score reliability. As
only 30% of score variance heterogeneity could be removed, this would
imply that differences in error score variance $\sigma^2_E$ were
responsible for less than a third of the heterogeneity in score
variances found. On the contrary, more than two thirds of heterogeneity
could be attributed to actual differences in how the underlying true
scores are distributed across samples.

Concerning the re-arranged inequality from Appendix A, we identify one
crucial condition or rule described in @eq-ineq_short.

@@

As long as $R_1^3$ is smaller or equal compared to $R_2$, @eq-ineq_init
is bound to hold true. This means that as long as the average score
reliability, to the power of 3, does not exceed the relative remaining
variance heterogeneity after the attenuation correction, @eq-ineq_init
is not violated. For example, assuming the average score reliability is
about .8, as long as $R_2$ is at least .512 or larger, the inequality
holds true and ES heterogeneity grows larger as a consequence of the
attenuation correction procedure. In such cases, the claims made in W&D
and H&S could not hold.

$R_1$ essentially describes the relative size of mean score variance
after the attenuation correction, and $R_2$ essentially described the
relative size of score variance heterogeneity after the attenuation
correction. Therefore, @eq-ineq_init states that for such a correction
procedure to truly reduce ES heterogeneity, the reduction in the
heterogeneity of score variance needs to be substantially larger than
the reduction in its mean values. The mean value reduction carries
stronger weight, as $R_1$ is taken to the power of 3 in @eq-ineq_init.
The smaller $R_1$ or $R_2$, the larger the reduction in heterogeneity or
mean value.

@@

It is important to note that the condition in @eq-ineq_short only
describes under which conditions the inequality is guaranteed to hold,
no matter what value other parameters like the mean value or
heterogeneity in mean differences take on. This means that, even if the
condition in @eq-ineq_short is violated, it is by no means guaranteed
that an attenuation correction procedure would actually reduce
heterogeneity in ES. This means that inference made from @eq-ineq_short
is asymmetrical. While we know, if the inequality holds, heterogeneity
in corrected ES will be larger than heterogeneity in uncorrected ES, we
can't be certain that the opposite is true if the inequality does not
hold.

@@

Estimation of $R_1$ and $R_2$ require estimates of meta-analytic mean
and heterogeneity of both the total score variance $\sigma^2_X$ and the
true score variance $\sigma^2_T$. Elsewhere, we proposed the method
boot-err (ref.) as an alternative method to test for and estimate the
differences in error score variance $\sigma^2_E$. In Beinhauer et al.
(ref.), we deliver detailed instructions on how estimates for mean and
heterogeneity of total score variance, and the score variance components
can be derived. For the total score variance $\sigma^2_X$, all we need
to do is to appropriately transform the observed estimates using the
variance-stabilizing properties of the natural log. Subsequently, we use
bootstrapping to derive robust standard errors for these transformed
observations. Using a random-efffects meta-analysis, we derive estimates
of mean and heterogeneity, which need to be backtransformed to the
original scale. Appropriate equation to back-transform estimates can be
found in Appendix B.

@@

For one of the phenomena, however, we observe a large disparity compared
to the rest of the results. For the Shnabel_Willingness_Reconcile_RPP
phenomenon, $R_2$ is very low, below .25. This makes it lower than its
$R_1^3 = .574$. This would imply that for this phenomenon, the
conditions for a reduction in ES heterogeneity would be met. However, in
@tbl-tauES, we find no such reduction. What becomes apparent, however,
is that for this phenomenon we are working with the low number of 9
administration sites. Estimation precision for heterogeneity in variance
components, even more so than for ES, largely depends on the sample
size. In Beinhauer et al. (ref) we come to the conclusion that at 10
items, which the Willingness to Reconcile scale consists of, a sample
size around 33 is required to achieve sufficient power for a statistical
test of score variance heterogeneity. Most likely, the low sample size
led to implausible estimates in both true and total score variance
heterogeneity and subsequently to an implausible estimate of $R_2$.
Meta-analytic estimates, especially those concerning score variance
components such as $R_2$, therefore should be interpreted with caution,
if the sample size was particularly low.

@@

Generally, @tbl-tauES implies that for almost all measurements
concerning these phenomena, the heterogeneity in observed score
variances could not be substantially explained by differences in
measurement quality. Only for one phenomenon, where differences in score
variance could be found, $R_2$ was smaller than $R_1^3$. As discussed,
for this phenomenon, the number of administration sites was too low to
accurately estimate heterogeneity in score variance components. For all
remaining phenomena, $R_2$ was larger than $R_1$, implying that an
attenuation correction procedure could not reduce ES heterogeneity. The
results in @tbl-tauES agree with this conclusion, as no phenomena could
be identified where statistically significant ES heterogeneity was
accompanied by a reduction in differences due to the correction.

```{r, echo = FALSE}
#| label: tbl-R1R2
#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
             col.names = c("MASC", 
                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
                           "$\\hat{\\mu}_{\\sigma^2_T}$", 
                           "$\\hat{\\tau}_{\\sigma^2_X}$", 
                           "$\\hat{\\tau}_{\\sigma^2_T}$", 
                           "$R_1$", 
                           "$R_2$"))
```

@@

In the examples we re-analysed throughout this text, as well as the
majority of research studies done in the behavioural sciences, no proper
sampling techniques are employed. Most samples where these instruments
are applied come to be through convenience sampling, where students
self-select into studies. Without a valid sampling scheme, we
essentially have have no basis to assume anything about the true score
variance. Most importantly, however, we have no reason to assume that
the true score variance could potentially be homogeneous or stable
across administration sites. Instead, it is very likely that the
convenience sampling procedures produce samples that vary extremely
strongly in how they are put together, leading to large degrees of true
score variance heterogeneity. At the same time, score reliability is a
low-hanging fruit to optimise as most researchers are aware and trained
in assessing score reliability.

@@

If we, in the space of Psychology, keep making use of convenience
samples simply because they are so convenient, we can not expect
anything to change. Most likely, ES heterogeneity will remain a
substantial factor hampering the interpretability of original and
replication studies. We could learn more about standardized effect sizes
and why our phenomena vary across settings, populations or continents by
not following the current status quo of convenience samples. Until this
status quo changes, however, we will keep wasting valuable resources by
attempts of discussing ES heterogeneity or assessing uninformative
differences in score reliability.

@@

The analytical arguments presented following Equation 17 rested on the
use of the delta-method, as Equation 17 contained parameters of the
score standard deviations’ distributions, while the $R_1$ and $R_2$
metrics were designed to discuss score variances. The delta-method rests
on a number of assumptions: (1) approximations derived from the
delta-method require for the transformed variable to follow a marginal
normal distribution (ref). Since the variable under discussion here is
the standard deviation, which is bounded to be larger than zero and has
a non-normal sampling distribution (ref), this assumption is most likely
violated; (2) the derivative of the transformation function needs to be
available. As the transformation function is essentially just the square
root, this assumption should hold up; (3) as the delta-method
approximates the parameters of the transformed variable using the
first-order Taylor-expansion, this implies that higher-order terms are
required to be negligible. In practice, this implies that the variance
of the untransformed variable needs to be small. However, as this entire
procedure discusses the influence of variance in standard deviations on
ES heterogeneity, introduced by differences in score reliability, this
assumption is probably violated as well. The violation of these
assumptions likely introduce a bias in estimates derived from these
approximations (ref). However, the inequality derived with use of the
delta-method is not used to actually derive any estimates, but to
discuss how an attenuation correction procedure can affect ES
heterogeneity. Therefore, we believe that the violation of these
assumptions is defensible in this case, as the analytical arguments
derived from the inequality should still hold up, nevertheless. \[add to
appendix!\] Additionally, we used the empirical examples available to
assess in how far estimates derived using the delta-method
approximations diverge from estimates derived without it. The
differences identified are rather miniscule at the third decimal,
reinforcing our belief that these assumption violations are defensible.
